Investigating the outer reaches of the Sculptor and Ursa Minor dwarf galaxies

### 2025-08-23 thesis/appendix.tex
thesis julio ch1 edits
thesis refining data section

Each likelihood is normalized over their respective 2D parameter space
for both the satellite. To control the relative frequency of member and
background stars, \(f_{\rm sat}\) representing the fraction of member
stars in the field. The total likelihood for any star in this model is
the sum of the satellite and background likelihoods, weighted by their
relative frequencies, \begin{equation}{
{\cal L}_{\rm tot} = f_{\rm sat}{\cal L}_{\rm sat} + (1-f_{\rm sat}){\cal L}_{\rm bg}.
}\end{equation} The probability that any star belongs to the satellite
is then given by \begin{equation}{
P_{\rm sat} = 
\frac{f_{\rm sat}\,{\cal L}_{\rm sat}}{{\cal L}_{\rm tot}}
= \frac{f_{\rm sat}{\cal L}_{\rm sat}}{f_{\rm sat}{\cal L}_{\rm sat} + (1-f_{\rm sat}){\cal L}_{\rm bg}}.
}\end{equation}

@@

\begin{figure}
\centering
\pandocbounded{\includegraphics[keepaspectratio]{/Users/daniel/thesis_2/figures/analytic_profile_comparison.pdf}}
\caption[Comparison of analytic density profiles]{A comparison of
different parameterizations for dwarf galaxy density profiles. Note that
deviations between profiles only become apparent past 3 R\_h, and only
the Plummer profile, in contrast to more commonly assumed profiles,
deviates by \textasciitilde1 dex positive before 6 Rh. Since this
profile is a far minority in the literature, deviations from exponential
and close relatives are interesting and worth further consideration.}
\end{figure}

>>>>

Each likelihood is normalized over their respective 2D parameter space
for both the satellite. To control the relative frequency of member and
background stars, \(f_{\rm sat}\) representing the fraction of member
stars in the field. The total likelihood for any star in this model is
the sum of the satellite and background likelihoods, weighted by their
relative frequencies,
\begin{equation}\protect\phantomsection\label{eq:Ltot}{
{\cal L}_{\rm tot} = f_{\rm sat}{\cal L}_{\rm sat} + (1-f_{\rm sat}){\cal L}_{\rm bg}.
}\end{equation} The probability that any star belongs to the satellite
is then given by \begin{equation}{
P_{\rm sat} = 
\frac{f_{\rm sat}\,{\cal L}_{\rm sat}}{{\cal L}_{\rm tot}}
= \frac{f_{\rm sat}{\cal L}_{\rm sat}}{f_{\rm sat}{\cal L}_{\rm sat} + (1-f_{\rm sat}){\cal L}_{\rm bg}}.
}\end{equation}

@@

\begin{figure}
\centering
\pandocbounded{\includegraphics[keepaspectratio]{figures/analytic_profile_comparison.pdf}}
\caption[Comparison of analytic density profiles]{A comparison of
different parameterizations for dwarf galaxy density profiles. Note that
deviations between profiles only become apparent past 3 R\_h, and only
the Plummer profile, in contrast to more commonly assumed profiles,
deviates by \textasciitilde1 dex positive before 6 Rh. Since this
profile is a far minority in the literature, deviations from exponential
and close relatives are interesting and worth further consideration.}
\end{figure}

### 2025-08-23 thesis/introduction.tex
thesis refining data section

Simulating dwarf galaxies accurately in a cosmological context remains a
substantial challenge. Currently, cosmological simulations can predict
the overall abundance of larger mass? dwarf galaxies {[}e.g., {]} and
broadly predict effects of tides \citep[e.g.,][]{riley+2024}. But, dwarf
galaxies are often near the resolution limit. Insufficient resolution
can lead to artificial disruption of dwarf galaxies and overestimation
of tidal effects \citep[e.g.,][]{santos-santos+2025}. For example, the
highest resolution cosmological simulation of a Milky Way-size dark
matter halo, the Aquarius project \citep{springel+2008}, achieved a DM
resolution of \(1.712 \times10^3 \Mo\), enough to barely resolve the
stellar components of Sculptor like halos (see chapter
{[}@-sec:methods{]}). To address this challenge, idealized simulations
only simulate a single subhalo in an approximate host potential. As a
result, idealized simulations can reach excellent numerical convergence
at the cost of realism. For instance, our simulations later are about 3x
higher resolution than Aquarius at a fraction of the computational cost
(400x less particles). Idealized simulations make numerous
simplifications: neglecting mergers, cosmological context and evolution,
mass assembly, and often baryonic physics. We use idealized simulations
here which are more powerful in accurately assessing tidal effects after
infall, given that the idealizations do not impact these galaxies's
recent history too much.

@@

In chapter {[}@-sec:observations{]}, we describe how we compute
observational density profiles from \citet{jensen+2024}. In chapter
{[}@-sec:methods{]}, we review our simulation methods. Next, we present
our results for the tidal effects on Sculptor and Ursa Minor in chapter
{[}@-sec:results{]}, We discuss our results, limitations, and
implications in chapter {[}@-sec:discussion{]}. Finally, chapter
{[}@-sec:summary{]} summarizes this work and discuss future directions
for similar work and the field of dwarf galaxies.

>>>>

Simulating dwarf galaxies accurately in a cosmological context remains a
substantial challenge. Currently, cosmological simulations can predict
the overall abundance of larger mass? dwarf galaxies {[}e.g., {]} and
broadly predict effects of tides \citep[e.g.,][]{riley+2024}. But, dwarf
galaxies are often near the resolution limit. Insufficient resolution
can lead to artificial disruption of dwarf galaxies and overestimation
of tidal effects \citep[e.g.,][]{santos-santos+2025}. For example, the
highest resolution cosmological simulation of a Milky Way-size dark
matter halo, the Aquarius project \citep{springel+2008}, achieved a DM
resolution of \(1.712 \times10^3 \Mo\), enough to barely resolve the
stellar components of Sculptor like halos (see chapter
\ref{sec:methods}). To address this challenge, idealized simulations
only simulate a single subhalo in an approximate host potential. As a
result, idealized simulations can reach excellent numerical convergence
at the cost of realism. For instance, our simulations later are about 3x
higher resolution than Aquarius at a fraction of the computational cost
(400x less particles). Idealized simulations make numerous
simplifications: neglecting mergers, cosmological context and evolution,
mass assembly, and often baryonic physics. We use idealized simulations
here which are more powerful in accurately assessing tidal effects after
infall, given that the idealizations do not impact these galaxies's
recent history too much.

@@

In chapter \ref{sec:observations}, we describe how we compute
observational density profiles from \citet{jensen+2024}. In chapter
\ref{sec:methods}, we review our simulation methods. Next, we present
our results for the tidal effects on Sculptor and Ursa Minor in chapter
\ref{sec:results}, We discuss our results, limitations, and implications
in chapter \ref{sec:discussion}. Finally, chapter \ref{sec:summary}
summarizes this work and discuss future directions for similar work and
the field of dwarf galaxies.
